VIPRE: Virtualized Intermittently-Powered Runtime Environment

### 2017-07-06 paper/intro.tex
put Brandon's propose section structure in the paper; initial corrections in Brandon's intro, added all author names in comments

Advances in processor efficiency along with the development of
energy-harvesting power systems has created a new category of devices that
require neither a battery nor a tethered power supply.  These devices operate
entirely using energy extracted from their environment, such as RF energy,
photovoltaics, and vibration.  Incorporating compute, storage, sensing, and
communication hardware, such energy-independent devices are a promising
candidate as the underlying technology for emerging internet of things (IoT)
applications, in- and on-body medical applications, and emerging applications,
like energy-harvesting nano-satellites.

While promising, energy-harvesting devices present a unique design challenge in
that they operate only {\em intermittently} when energy is available.   An
energy-harvesting device buffers energy from its environment, e.g., in a
capacitor. When a threshold quantity of energy accumulates, the device begins
operating.  Harvestable energy sources typically produce relatively very little
power compared to a platform's operating power level.  A devices operates for a
brief period of time and when buffered energy is depeleted, shuts down and
begins recharging to later operate again.  Charge and discharge times vary by
device and some devices~\cite{wisp} may fail 10 to 100 times per second.

Software in an energy-harvesting system operates according to the {\em
intermittent execution model}~\cite{dino}, with bursts of execution that are
interrupted by failure periods. When power fails, a device loses its volatile
state (i.e., registers, stack, SRAM) and retains its non-volatile state (i.e.,
FRAM, Flash). While capturing periodic checkpoints~\cite{mementos,tictpl,quickrecall}
and sleep scheduling~\cite{dewdrop,hibernus,hibernusplusplus} help to preserve
execution progress, failures can leave volatile and non-volatile state inconsistent,
leading to unrecoverable failures~\cite{mspcdino,edb,edbtoppicks}.

@@

%FRAM costs more than FRAM -- need to figure out how to execute out of SRAM
New programming and execution models~\cite{dino,ratchet,chain,alpaca} and
architecture mechanisms~\cite{clank, idetic, nvp} help avoid data
inconsistency.  New architectures require costly hardware changes and are
inapplicable to today's systems.  New programming models and compilers force
programmers to use new memory models~\cite{chain,ratchet}, which may inhibit
adoption.  New memory models may also increase the amount of non-volatile
memory accesses by requring non-volatile versioning~\cite{chain} or completely
precluding the use of volatile memory~\cite{ratchet}.  Increasing pressure on
non-volatile memory or precluding the use of volatile memory may degrade
efficiency; e.g., FRAM has a higher access energy and latency than
SRAM~\cite{alpaca,wisp}.  Commercially available devices are often provisioned
with much less SRAM than FRAM~\cite{wolverine,wisp} and effectively leveraging
the relatively high efficiency of SRAM is a key challenge for task-based
intermittent systems.

@@

We evaluated our system on a collection of benchmark programs taken from the
literature, and compare directly to prior work, showing that \sys has high
performance and can flexible target a variety of platforms without recompiling
the code.  We further demonstrate the efficiency and versatility of \sys with
an end-to-end evaluation of a sensing and data processing application deployed
in a lab environment \TODO{blah blah blah eval}.

>>>>

Advances in processor efficiency along with the development of energy-harvesting power systems has created a new category of devices that require neither a battery nor a tethered power supply. These devices operate entirely using energy extracted from their environment, such as RF energy~\cite{}, photovoltaics~\cite{}, and vibration~\cite{}. Incorporating compute, storage, sensing, and communication hardware, such energy-independent devices are a promising candidate as the underlying technology for emerging Internet of Things (IoT) applications, in- and on-body medical applications~\cite{}, and emerging applications, like energy-harvesting nano-satellites~\cite{}.

While promising, energy-harvesting devices present a unique design challenge in
that they operate only {\em intermittently} when energy is available. An
energy-harvesting device buffers energy from its environment, e.g., in a
capacitor. When a threshold quantity of energy accumulates, the device begins
operating. Harvestable energy sources typically produce relatively very little
power compared to a platform's operating power level. A devices operates for a
brief period of time and when buffered energy is depleted, shuts down and
begins recharging to later operate again. Charge and discharge times vary by
device and some devices~\cite{wisp} may fail 10 to 100 times per second~\cite{}.

Software in an energy-harvesting system operates according to the {\em
intermittent execution model}~\cite{dino}, with bursts of execution that are
interrupted by failure periods. When power fails, a device loses its volatile
state, i.e., registers, stack, SRAM, and retains its non-volatile state, i.e.,
FRAM, Flash. While capturing periodic checkpoints~\cite{mementos,tictpl,quickrecall}
and sleep scheduling~\cite{dewdrop,hibernus,hibernusplusplus} help to preserve
execution progress, failures can leave volatile and non-volatile state inconsistent, leading to unrecoverable failures~\cite{mspcdino,edb,edbtoppicks}.

@@

%FRAM costs more than FRAM -- need to figure out how to execute out of SRAM
New programming and execution models~\cite{dino,ratchet,chain,alpaca} and
architecture mechanisms~\cite{clank,idetic,nvp} help avoid data
inconsistency. New architectures require costly hardware changes and are
inapplicable to today's systems. New programming models and compilers force
programmers to use new memory models~\cite{chain,ratchet}, which may inhibit
adoption. New memory models may also increase the amount of non-volatile
memory accesses by requiring non-volatile versioning~\cite{chain} or completely
precluding the use of volatile memory~\cite{ratchet}. Increasing pressure on
non-volatile memory or precluding the use of volatile memory may degrade
efficiency; e.g., FRAM has a higher access energy and latency than
SRAM~\cite{alpaca,wisp}. Commercially available devices are often provisioned
with much less SRAM than FRAM~\cite{wolverine,wisp} and effectively leveraging
the relatively high efficiency of SRAM is a key challenge for task-based
intermittent systems.

@@

We evaluated our system on a collection of benchmark programs taken from the
literature, and compare directly to prior work, showing that \sys has high
performance and can flexible target a variety of platforms without recompiling
the code. We further demonstrate the efficiency and versatility of \sys with
an end-to-end evaluation of a sensing and data processing application deployed
in a lab environment. \TODO{blah blah blah eval}.
